---
editor_options: 
   chunk_output_type: console
---

```{r}
#| label: setup-common-01
#| include: false
#| warning: false
source("includes/_common.R")
# ACA FUL
```

# ACA Federal Upper Limits {#sec-acaful}

__Affordable Care Act Federal Upper Limits (FUL)__ based on the weighted average of the most recently reported monthly __average manufacturer price (AMP)__ for pharmaceutical and therapeutically equivalent multiple source drug products that are available for purchase by retail community pharmacies on a nationwide basis.

## Metadata

```{r}
#| label: s7_object
#| echo: false
end <- caid_endpoint("aca_ful")
print_meta(end)
```

### Dictionary

```{r}
#| label: dict
#| echo: false
#| message: true
caid_dictionary("Federal Upper Limits Data Dictionary") |> 
  print_dict_tbl()
```

## Data

```{r}
#| label: http_request
#| message: false
#| warning: false
#| echo: false
resp <- quick("aca_ful", offset = 0, limit = 8000) |>
  mtt(
    mdr_unit_type       = factor_(mdr_unit_type),
    route               = factor_(route),
    amp_wt_avg          = as.double(weighted_average_of_amps),
    aca_ful             = as.double(aca_ful),
    package_size        = as.integer(package_size),
    year                = as.integer(year),
    month               = as.integer(month),
    arated              = charbin(arated),
    gt_175pc_amp_wt_avg = charbin(
      multiplier_greater_than_175_percent_of_weighted_avg_of_amps
    ),
    multiplier_greater_than_175_percent_of_weighted_avg_of_amps = NULL,
    weighted_average_of_amps = NULL
  )

resp

resp |> cheapr::overview()
```

### Distribution

```{r}
#| label: distribution
#| message: false
#| warning: false
#| echo: false
res <- describe_distribution(resp)
res
```

```{r}
#| label: distribution_plot
#| layout-ncol: 2
#| message: false
#| warning: false
#| echo: false
plots <- plot(res, dispersion = TRUE)
walk(plots, print)
```
