---
editor_options: 
   chunk_output_type: console
---

```{r}
#| label: setup-common-01
#| include: false
source("includes/_common.R")

unscore <- \(x) gsub("___owner$", "", x, perl = TRUE)
charbin <- \(x) val_match(x, "N" ~ 0L, "Y" ~ 1L)
as_prop <- \(x) case(x == "0" ~ 0, is_na(x) ~ NA_real_, .default = as.double(x) / 100)

rhc_names <- c(
    "ENROLLMENT ID"                = "enid",
    "ENROLLMENT STATE"             = "enid_state",
    "PROVIDER TYPE CODE"           = "prov_code",
    "PROVIDER TYPE TEXT"           = "prov_desc",
    "NPI"                          = "npi",
    "MULTIPLE NPI FLAG"            = "multi_npi",
    "CCN"                          = "ccn",
    "ASSOCIATE ID"                 = "pac",
    "ORGANIZATION NAME"            = "org_name",
    "DOING BUSINESS AS NAME"       = "dba_name",
    "INCORPORATION DATE"           = "inc_date",
    "INCORPORATION STATE"          = "inc_state",
    "ORGANIZATION TYPE STRUCTURE"  = "org_type",
    "ORGANIZATION OTHER TYPE TEXT" = "org_other",
    "PROPRIETARY_NONPROFIT"        = "prop_non",
    "ADDRESS LINE 1"               = "address_line_1",
    "ADDRESS LINE 2"               = "address_line_2",
    "CITY"                         = "city",
    "STATE"                        = "state",
    "ZIP CODE"                     = "zip"
  )

rhc_owner_names <- c(
  "ENROLLMENT ID"                       = "enid",
  "ASSOCIATE ID"                        = "pac",
  "ORGANIZATION NAME"                   = "org_name",
  "ASSOCIATE ID - OWNER"                = "own_pac",
  "TYPE - OWNER"                        = "own_type",
  "ROLE CODE - OWNER"                   = "own_role_cd",
  "ROLE TEXT - OWNER"                   = "own_role_desc",
  "ASSOCIATION DATE - OWNER"            = "own_asoc_date",
  "FIRST NAME - OWNER"                  = "own_first_name",
  "MIDDLE NAME - OWNER"                 = "own_middle_name",
  "LAST NAME - OWNER"                   = "own_last_name",
  "TITLE - OWNER"                       = "own_title",
  "ORGANIZATION NAME - OWNER"           = "own_org_name",
  "DOING BUSINESS AS NAME - OWNER"      = "own_dba_name",
  "ADDRESS LINE 1 - OWNER"              = "own_address_line_1",
  "ADDRESS LINE 2 - OWNER"              = "own_address_line_2",
  "CITY - OWNER"                        = "own_city",
  "STATE - OWNER"                       = "own_state",
  "ZIP CODE - OWNER"                    = "own_zip",
  "PERCENTAGE OWNERSHIP"                = "own_pct_owned",
  "CREATED FOR ACQUISITION - OWNER"     = "is_acq",
  "CORPORATION - OWNER"                 = "is_corp",
  "LLC - OWNER"                         = "is_llc",
  "MEDICAL PROVIDER SUPPLIER - OWNER"   = "is_supp",
  "MANAGEMENT SERVICES COMPANY - OWNER" = "is_mgmt",
  "MEDICAL STAFFING COMPANY - OWNER"    = "is_staff",
  "HOLDING COMPANY - OWNER"             = "is_hold",
  "INVESTMENT FIRM - OWNER"             = "is_firm",
  "FINANCIAL INSTITUTION - OWNER"       = "is_bank",
  "CONSULTING FIRM - OWNER"             = "is_consult",
  "FOR PROFIT - OWNER"                  = "is_forprof",
  "NON PROFIT - OWNER"                  = "is_nonprof",
  "OTHER TYPE - OWNER"                  = "own_other_type",
  "OTHER TYPE TEXT - OWNER"             = "own_other_text"
)

owner_role <- list(
  "01" = "5% OR MORE OWNERSHIP INTEREST",
  "03" = "PARTNER",
  "25" = "CONTRACTED MANAGING EMPLOYEE",
  "34" = "5% OR GREATER DIRECT OWNERSHIP INTEREST",
  "35" = "5% OR GREATER INDIRECT OWNERSHIP INTEREST",
  "36" = "5% OR GREATER MORTGAGE INTEREST",
  "37" = "5% OR GREATER SECURITY INTEREST",
  "38" = "GENERAL PARTNERSHIP INTEREST",
  "39" = "LIMITED PARTNERSHIP INTEREST",
  "40" = "OFFICER",
  "41" = "DIRECTOR",
  "42" = "W-2 MANAGING EMPLOYEE",
  "43" = "OPERATIONAL/MANAGERIAL CONTROL",
  "44" = "OTHER"
)
```

# Rural Health Clinics {#sec-rhc}

## RHC Enrollments

The __Rural Health Clinic (RHC) Enrollments__ dataset provides enrollment information on all RHCs currently enrolled in Medicare. This data includes information on the RHCs legal business name, doing business as name, organization type and address.

### Metadata

```{r}
#| label: s7_rhc_enroll
#| echo: false
end  <- care_endpoint("rhc_enrollments")
print_meta(end, "care")
```

### Resources

```{r}
#| label: resources_rhc_enroll
#| echo: false
print_resources(end)
```

### Dictionary

```{r}
#| label: dict_rhc_enroll
#| echo: false
dict_rhc_enroll <- list(
  "ENROLLMENT ID"                = "RHC's enrollment ID.",
  "ENROLLMENT STATE"             = "RHC's enrollment state.",
  "PROVIDER TYPE CODE"           = "RHC's provider type code.",
  "PROVIDER TYPE TEXT"           = "RHC's provider type description.",
  "NPI"                          = "RHC's NPI.",
  "MULTIPLE NPI FLAG"            = "Indicates if RHC has multiple NPIs.",
  "CCN"                          = "RHC's CMS Certification Number (CCN), formerly called an OSCAR Number.",
  "ASSOCIATE ID"                 = "RHC's PECOS Associate Control (PAC) ID.",
  "ORGANIZATION NAME"            = "RHC's legal business name.",
  "DOING BUSINESS AS NAME"       = "RHC's doing-business-as name.",
  "INCORPORATION DATE"           = "Date RHC was incorporated.",
  "INCORPORATION STATE"          = "State RHC was incorporated.",
  "ORGANIZATION TYPE STRUCTURE"  = "RHC's organization type structure.",
  "ORGANIZATION OTHER TYPE TEXT" = "Description of the organization structure if OTHER.",
  "PROPRIETARY_NONPROFIT"        = "P if registered as proprietor with the IRS; N if registered as non-profit.",
  "ADDRESS LINE 1"               = "Address line 1 of the RHC's practice location address.",
  "ADDRESS LINE 2"               = "Address line 1 of the RHC's practice location address.",
  "CITY"                         = "City of the RHC's practice location address.",
  "STATE"                        = "State of the RHC's practice location address.",
  "ZIP CODE"                     = "Zip code of the RHC's practice location address."
)
print_dict_list(dict_rhc_enroll)
```

### Data

```{r}
#| label: http_request_rhc_enroll
#| message: true
#| echo: false
enroll_resp <- list(
  quick("rhc_enrollments", offset = 0, limit = 5000),
  quick("rhc_enrollments", offset = 5000, limit = 5000)) |> 
  list_rbind() |>
  rnm(rhc_names) |>
  mtt(inc_date = providertwo:::as_date(inc_date),
      multi_npi = charbin(multi_npi),
      address = providertwo:::make_address(address_line_1, address_line_2),
      org_type = ifelse(org_type == "OTHER", org_other, org_type)) |>
  slt(-address_line_1, -address_line_2, -org_other)

enroll_resp |> data_codebook()
```

## RHC All Owners

The __Rural Health Clinic (RHC) All Owners__ dataset provides ownership information on all RHCs currently enrolled in Medicare. This data includes ownership information such as ownership name, ownership type, ownership address and ownership effective date.

### Metadata

```{r}
#| label: s7_rhc_owners
#| echo: false
owners  <- care_endpoint("rhc_owners")
print_meta(owners, "care")
```

### Resources

```{r}
#| label: resources_rhc_owners
#| echo: false
print_resources(owners)
```

### Dictionary

```{r}
#| label: dict_rhc_owners
#| echo: false
dict_rhc_owners <- list(
  "ENROLLMENT ID"                       = "RHC's enrollment ID.",
  "ASSOCIATE ID"                        = "RHC's PECOS Associate Control (PAC) ID.",
  "ORGANIZATION NAME"                   = "RHC's legal business name.",
  "ASSOCIATE ID - OWNER"                = "PAC ID of the Owner; an associate with ownership or managing control interest in the RHC enrollment.",
  "TYPE - OWNER"                        = "Type of owner, I for individual and O for organization.",
  "ROLE CODE - OWNER"                   = "Owner's role code.",
  "ROLE TEXT - OWNER"                   = "Description for Owner Role Code.",
  "ASSOCIATION DATE - OWNER"            = "Date owner became associated with the RHC.",
  "FIRST NAME - OWNER"                  = "Individual owner's first name.",
  "MIDDLE NAME - OWNER"                 = "Individual owner's middle name.",
  "LAST NAME - OWNER"                   = "Individual owner's last name.",
  "TITLE - OWNER"                       = "Individual owner's title.",
  "ORGANIZATION NAME - OWNER"           = "Organization owner's legal business name.",
  "DOING BUSINESS AS NAME - OWNER"      = "Organization owner's doing-business-as name.",
  "ADDRESS LINE 1 - OWNER"              = "Address line 1 of organization owner's mailing address.",
  "ADDRESS LINE 2 - OWNER"              = "Address line 2 of organization owner's mailing address.",
  "CITY - OWNER"                        = "City of organization owner's mailing address.",
  "STATE - OWNER"                       = "State of organization owner's mailing address.",
  "ZIP CODE - OWNER"                    = "Zip code of organization owner's mailing address.",
  "PERCENTAGE OWNERSHIP"                = "Owner's ownership percentage in the RHC enrollment.",
  "CREATED FOR ACQUISITION - OWNER"     = "Organization owner was solely created to acquire the RHC and/or it's assets.",
  "CORPORATION – OWNER"                 = "Owner's organization type is corporation.",
  "LLC – OWNER"                         = "Owner's organization type is limited liability company.",
  "MEDICAL PROVIDER SUPPLIER – OWNER"   = "Owner's organization type is medical provider/supplier.",
  "MANAGEMENT SERVICES COMPANY – OWNER" = "Owner's organization type is management services company.",
  "MEDICAL STAFFING COMPANY – OWNER"    = "Owner's organization type is medical staffing company.",
  "HOLDING COMPANY – OWNER"             = "Owner's organization type is holding company.",
  "INVESTMENT FIRM – OWNER"             = "Owner's organization type is investment firm.",
  "FINANCIAL INSTITUTION – OWNER"       = "Owner's organization type is bank or other financial institution.",
  "CONSULTING FIRM – OWNER"             = "Owner's organization type is consulting firm.",
  "FOR PROFIT – OWNER"                  = "Owner's organization type is for profit.",
  "NON PROFIT – OWNER"                  = "Owner's organization type is non-profit.",
  "OTHER TYPE – OWNER"                  = "Owner's organization type is not listed on the CMS form.",
  "OTHER TYPE TEXT – OWNER"             = "Other type of owner's organization that is not listed on the CMS form."
)
print_dict_list(dict_rhc_owners)
```

### Data

```{r}
#| label: http_request_rhc_owners
#| message: true
#| echo: false
owners_resp <- list(
  quick("rhc_owners", offset = 0, limit = 5000),
  quick("rhc_owners", offset = 5000, limit = 5000)) |> 
  list_rbind() |>
  rnm(rhc_owner_names) |>
  mtt(own_asoc_date = providertwo:::as_date(own_asoc_date),
      own_pct_owned = as_prop(own_pct_owned),
      own_address = providertwo:::make_address(own_address_line_1, own_address_line_2),
      own_role = glue("[{own_role_cd}] {own_role_desc}"),
      own_type = factor_(own_type),
      own_other_type = charbin(own_other_type),
      is_acq = charbin(is_acq),
      is_corp = charbin(is_corp),
      is_llc = charbin(is_llc),
      is_supp = charbin(is_supp),
      is_mgmt = charbin(is_mgmt),
      is_staff = charbin(is_staff),
      is_hold = charbin(is_hold),
      is_firm = charbin(is_firm),
      is_bank = charbin(is_bank),
      is_consult = charbin(is_consult),
      is_forprof = charbin(is_forprof),
      is_nonprof = charbin(is_nonprof)
      ) |>
  slt(-own_address_line_1, 
      -own_address_line_2,
      -own_role_cd,
      -own_role_desc)

owners_resp |> data_codebook()
```


## Join

```{r}
#| label: join_rhc
#| message: false
#| echo: false
join(
  owners_resp,
  enroll_resp,
  on = c("enid", "pac", "org_name"),
  overid = 2
) |> 
  str()
```

