---
editor_options: 
   chunk_output_type: console
---

```{r}
#| label: setup-common-01
#| include: false
#| warning: false
source("includes/_common.R")

enrollees = c(
  NPI                = "npi",
  PECOS_ASCT_CNTL_ID = "pac",
  ENRLMT_ID          = "enid",
  PROVIDER_TYPE_CD   = "specialty_code",
  PROVIDER_TYPE_DESC = "specialty_desc",
  STATE_CD           = "state",
  FIRST_NAME         = "first_name",
  MDL_NAME           = "middle_name",
  LAST_NAME          = "last_name",
  ORG_NAME           = "org_name"
)
```

# Provider Enrollment {#sec-enrollees}

The __Medicare Fee-For-Service Public Provider Enrollment__ dataset includes information on providers who are actively approved to bill Medicare or have completed the 855O at the time the data was pulled from the _Provider Enrollment, Chain, and Ownership System (PECOS)_. The release of this provider enrollment data is not related to other provider information releases such as Physician Compare or Data Transparency.

### Metadata

```{r}
#| label: s7_object
#| echo: false
end  <- care_endpoint("care_enrollees")
print_meta(end, "care")
```

### Resources

```{r}
#| label: resources
#| echo: false
print_resources(end)
```

## Data

```{r}
#| label: http_request
#| message: true
#| echo: false
resp <- quick("care_enrollees", offset = 5000, limit = 5000) |> 
  rnm(enrollees)
resp |> data_codebook()
```
