---
editor_options: 
   chunk_output_type: console
---

```{r}
#| label: setup-common-01
#| include: false
#| warning: false
source("includes/_common.R")
```

# CPC Initiative {#care-cpc}

CPC Initiative Participating Primary Care Practices

## Description 

The Comprehensive Primary Care (CPC) Initiative - Participating Primary Care Practices dataset provides a list of practices involved in a multi-payer initiative which fosters collaboration between public and private health care payers to strengthen primary care.

### Metadata

```{r}
#| label: s7_object
#| echo: false
care_endpoint("cpc_primary")
```

### Resources

```{r}
#| label: resources
#| echo: false
care_endpoint("cpc_primary") |> list_resources()
```

### Dictionary

```{r}
#| label: data_dict
#| echo: false
cpc_init_dict <- new_tbl(
  FIELD = c(
    "Name of Initiative", 
    "Participating Practice", 
    "Participating Practice Location", 
    "State", 
    "City", 
    "Geographic Reach", 
    "Street Address", 
    "Zip Code"
  ),
  DESCRIPTION = c(
    "Unencrypted ACO Identifier. This identifier can be linked to the encrypted ACO identifier used for prior performance years (PY) using the ACO ID Crosswalk available at https://www.cms.gov/Medicare/Medicare-Fee-for-ServicePayment/sharedsavingsprogram/program-data.", 
    "ACO Doing Business As (DBA) or Legal Business Name (LBN). Listed name is DBA unless DBA is not available, in which case LBN is used.", 
    "Indicates ACO description of use of advance investment payments.", 
    "Indicates which one of the three required categories the payment use falls under: increased staffing, health care infrastructure, or the provision of accountable care for underserved beneficiaries.", 
    "Indicates a selected general spend subcategory under a general spend category that a payment use falls under.", 
    "The dollar amount of AIP funding the ACO has received from CMS through December 2024.", 
    "The dollar amount the ACO estimates will be spent on the payment use during the 2024 performance year.", 
    "The dollar amount the ACO actually spent on the payment use during the 2024 performance year.", 
    "The dollar amount the ACO estimates will be spent on the payment use during the 2025 performance year.",
    "The dollar amount the ACO actually spent on the payment use during the 2025 performance year."
  )
)

cpc_init_dict |> 
  gt::gt_preview(20, 10) |>
  gtUtils::gt_theme_athletic() |> 
  gt::cols_align(align = "left", columns = gt::everything()) |> 
  gt::tab_options(
    quarto.disable_processing = TRUE,
    column_labels.font.weight = "bolder",
    column_labels.font.size = 14,
    table.font.size = 14,
    table.width = gt::pct(100)
  ) |> 
  gt::cols_align(align = "left", columns = gt::everything())
```


## Response

```{r}
#| label: http_request
#| message: true
resp <- quick("cpc_primary", limit = 5000)

resp |> 
  gt::gt_preview(20, 20) |>
  gtUtils::gt_theme_athletic() |>
  gt::tab_options(
    quarto.disable_processing = TRUE,
    column_labels.font.weight = "bolder",
    column_labels.font.size = 14,
    table.font.size = 14,
    table.width = gt::pct(100)
  ) |> 
  gt::cols_align(align = "left", columns = gt::everything())
```

## Overview

```{r}
#| label: datawizard_codebook
#| echo: false
resp |> 
  datawizard::data_codebook() |> 
  slt(-.row_id) |>
  gt::gt() |>
  gtUtils::gt_theme_athletic() |>
  gt::tab_options(
    quarto.disable_processing = TRUE,
    column_labels.font.weight = "bolder",
    column_labels.font.size = 16,
    column_labels.background.color = "black",
    column_labels.border.bottom.color = "white",
    table.font.size = 12,
    table.width = gt::pct(100)
  ) |>
  gt::cols_align(align = "left", columns = gt::everything()) |>
  gt::sub_missing(columns = gt::everything(), missing_text = "-")
```
