---
editor_options: 
   chunk_output_type: console
---

```{r}
#| label: setup-common-01
#| include: false
#| warning: false
source("includes/_common.R")
```

# DRA HAC Measures {#care-drahac}

The __Publicly Reported Deficit Reduction Act (DRA) Hospital-Acquired Condition (HAC) Measures__ data provides information on hospital-level measures rates for four of the HACs included in the DRA HAC payment provision for Medicare fee-for-service discharges:

   1. Foreign Object Retained After Surgery
   1. Blood Incompatibility
   1. Air Embolism
   1. Falls and Trauma

[The Publicly Reported DRA HAC Measures are for informational and quality improvement purposes only; the results do not affect payment and are not a part of the HAC Reduction Program.]{.aside}

## Metadata

```{r}
#| label: s7_object
#| echo: false
end <- care_endpoint("hac_measures")
# end@metadata$description
print_meta(end)
```

### Resources

```{r}
#| label: resources
#| echo: false
print_resources(end)
```

### Dictionary

```{r}
#| label: data_dict
#| echo: false
# browse_link(end, link = "dictionary")
new_tbl(
  field = c(
    "Provider_ID", 
    "Measure", 
    "Rate", 
    "Footnote", 
    "Start_Quarter", 
    "End_Quarter"
    ),
  description = c(
    "Provider's CCN ID (6-digit format).",
    "DRA HAC measure name. One of: Foreign Object Retained After Surgery, Air Embolism, Blood Incompatibility, or Falls and Trauma (includes fractures, dislocations, intracranial injuries, crushing injuries, burns, other injuries).",
    "Provider's DRA HAC measure rate result, for all four DRA HAC measures. All four DRA HAC measures' rates are reported as observed rates per 1,000 discharges.",
    "Provider's DRA HAC measure rate result's footnote. One of: 23 (Data based on claims submitted to CMS by facility, which has since reported discrepancies in their claims data.)",
    "First quarter of the DRA HAC measure's discharge period, for all four DRA HAC measures.",
    "Last quarter of the DRA HAC measure's discharge period, for all four DRA HAC measures."
  )
)
```

## Data

```{r}
#| label: http_request
#| message: false
#| warning: false
#| echo: false
resp <- list(
  quick("hac_measures", offset = 0, limit = 5000),
  quick("hac_measures", offset = 5000, limit = 5000),
  quick("hac_measures", offset = 10000, limit = 5000)) |> 
  purrr::list_rbind() |> 
  mtt(Measure = cheapr_if_else(
    Measure == "Falls and Trauma (Includes fractures, dislocations, intracranial injuries, crushing injuries, burns, other injuries)",
    "Falls and Trauma",
    Measure) |> factor_(),
      Rate = as.numeric(Rate)) |> 
  sbt(Rate > 0)
resp
```

## Distribution

```{r}
#| label: distribution
#| message: false
#| warning: false
#| echo: false
res <- describe_distribution(resp, by = "Measure")
res
```

```{r}
#| label: distribution_plot
#| message: false
#| warning: false
#| echo: false
plots <- describe_distribution(resp) |> 
  plot(dispersion = TRUE)
walk(plots, print)
```
