---
editor_options: 
   chunk_output_type: console
---

```{r}
#| label: setup-common-01
#| include: false
#| warning: false
source("includes/_common.R")
```

# DRA HAC Measures {#sec-drahac}

The __Publicly Reported Deficit Reduction Act (DRA) Hospital-Acquired Condition (HAC) Measures__ data provides information on hospital-level measures rates for four of the HACs included in the DRA HAC payment provision for Medicare fee-for-service discharges:

   1. Foreign Object Retained After Surgery
   1. Blood Incompatibility
   1. Air Embolism
   1. Falls and Trauma

[The Publicly Reported DRA HAC Measures are for informational and quality improvement purposes only; the results do not affect payment and are not a part of the HAC Reduction Program.]{.aside}

## Metadata

```{r}
#| label: s7_object
#| echo: false
end <- care_endpoint("hac_measures")
print_meta(end)
```

### Resources

```{r}
#| label: resources
#| echo: false
print_resources(end)
```

### Dictionary

```{r}
#| label: dict
#| echo: false
# browse_link(end, link = "dictionary")
dict <- new_tbl(
  field = c(
    "Provider_ID", 
    "Measure", 
    "Rate", 
    "Footnote", 
    "Start_Quarter", 
    "End_Quarter"
    ),
  description = c(
    "Provider's CCN.",
    "DRA HAC measure name: Foreign Object Retained After Surgery, Air Embolism, Blood Incompatibility, or Falls and Trauma (incl. fractures, dislocations, intracranial injuries, crushing injuries, burns, other injuries).",
    "DRA HAC measure rate result, reported as observed rates per 1,000 discharges.",
    "Rate footnote: Data based on facility claims submitted to CMS with reported discrepancies.",
    "First Quarter of the DRA HAC measure's discharge period.",
    "Last quarter of the DRA HAC measure's discharge period."
  )
)
print_dict_tbl(dict)
```

## Data

```{r}
#| label: http_request
#| message: false
#| warning: false
#| echo: false
resp <- list(
  quick("hac_measures", offset = 0, limit = 5000),
  quick("hac_measures", offset = 5000, limit = 5000),
  quick("hac_measures", offset = 10000, limit = 5000)) |> 
  purrr::list_rbind() |> 
  mtt(Measure = cheapr_if_else(
    Measure == "Falls and Trauma (Includes fractures, dislocations, intracranial injuries, crushing injuries, burns, other injuries)",
    "Falls and Trauma",
    Measure) |> factor_(),
      Rate = as.numeric(Rate)) |> 
  sbt(Rate > 0)
resp
```

## Distribution

```{r}
#| label: distribution
#| message: false
#| warning: false
#| echo: false
res <- describe_distribution(resp, by = "Measure")
res
```

```{r}
#| label: distribution_plot
#| message: false
#| warning: false
#| echo: false
plots <- describe_distribution(resp) |> 
  plot(dispersion = TRUE)
walk(plots, print)
```
