---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Format Experiments {#sec-bslib}

```{r}
#| label: setup
#| include: false
source("includes/_common.R")

knitr::opts_chunk$set(
  collapse  = FALSE,
  echo      = TRUE, 
  message   = TRUE, 
  warning   = FALSE,
  error     = TRUE,
  comment   = "",
  dpi       = 300, 
  out.width = "100%"
)
source("includes/_common.R")
library(bslib)
library(htmltools)
library(bsicons)
```

The __Agency for Healthcare Research and Quality (AHRQ) Patient Safety Indicator 11 (PSI-11) Measure Rates__ dataset provides information on provider-level measure rates regarding one preventable complication (_Postoperative Respiratory Failure_) for Medicare Fee-For-Service discharges. 

[The PSI-11 measure data are solely reported for providers' information and quality improvement purposes and are not a part of the __Deficit Reduction Act (DRA)__, __Hospital-Acquired Condition (HAC) Payment Provision__ or __HAC Reduction Program__.]{.aside}

## Metadata

```{r}
#| label: s7_object
#| echo: false
end  <- care_endpoint("ahqr_psi11")
print_meta(end)
end@metadata
end@dimensions@rows
```

### Value Boxes

```{r}
#| label: value_box
#| echo: false
#| output: "asis"
vbs <- list(
  value_box(
    title = "Rows",
    value = end@dimensions@rows,
    showcase = bs_icon("body-text")
  ),
  value_box(
    title = "Fields",
    value = end@dimensions@fields |> length(),
    showcase = bs_icon("boxes")
  ),
  value_box(
    title = "3rd value",
    value = "789",
    showcase = bs_icon("pie-chart"),
    p("The 4th detail"),
    p("The 5th detail"),
    p("The 6th detail")
  )
)

layout_column_wrap(
  width = "250px",
  !!!vbs
)
```

```{r}
layout_columns(
  value_box(
    title = "Lines of Code",
    value = "46,157",
    theme = "bg-gradient-green-teal",
    showcase = fontawesome::fa_i("box-open"),
    showcase_layout = "top right",
    full_screen = FALSE,
    fill = TRUE,
    height = NULL
  ),
  value_box(
    title = "Project budget variance",
    value = "$2,345.67",
    theme = "yellow",
    showcase = fontawesome::fa_i("creative-commons-share"),
    showcase_layout = "top right",
    full_screen = FALSE,
    fill = TRUE,
    height = NULL
  ),
  value_box(
    title = "Website loading time",
    value = "2.3 seconds",
    theme = "bg-gradient-indigo-purple",
    showcase = fontawesome::fa_i("i"),
    showcase_layout = "top right",
    full_screen = FALSE,
    fill = TRUE,
    height = NULL
  )
)
```

### Cards

```{r}
x <- card("A simple card")

page_fillable(
  fillable_mobile = TRUE,
  layout_columns(
    col_widths = breakpoints(
      sm = c(12, 12, 12),
      md = c(6, 6, 12),
      lg = c(4, 4, 4)
    ),
    x, x, x
  )
)
```

### Accordion

```{r}
items <- lapply(LETTERS[1:5], function(x) {
  accordion_panel(paste("Section", x), paste("Some narrative for section", x))
})

accordion(!!!items)
```


## Resources

```{r}
#| label: resources
#| echo: false
print_resources(end)
```

## Dictionary

```{r}
#| label: data_dict
#| echo: false
new_tbl(
  field = c(
    "HOSP_ID", 
    "ADM_DISC", 
    "RATE", 
    "INTERVAL_LOWER_LIMIT", 
    "INTERVAL_HIGHER_LIMIT", 
    "START_QUARTER", 
    "START_DATE", 
    "END_QUARTER", 
    "END_DATE"
  ),
  description = c(
    "Provider's CCN.", 
    "Provider's number of eligible discharges used for PSI-11 measure calculations.", 
    "Provider's PSI-11 smoothed measure rate (per 1k discharges).", 
    "95% confidence interval lower limit surrounding provider's smoothed measure rate.", 
    "95% confidence interval upper limit surrounding provider's smoothed measure rate.", 
    "First quarter of PSI-11 measure's performance period.", 
    "First date of PSI-11 measure's performance period.", 
    "Last quarter of PSI-11 measure's performance period.", 
    "Last date of PSI-11 measure's performance period."
  )
) |> 
  print_dict_tbl()
```

## Data

```{r}
#| label: http_request
#| message: true
#| echo: false
resp <- quick("ahqr_psi11", offset = 0, limit = 5000) |> 
  mtt(START_DATE           = providertwo:::as_date(START_DATE, fmt = "%m/%d/%Y"),
      END_DATE             = providertwo:::as_date(END_DATE, fmt = "%m/%d/%Y"),
      ADM_DISC             = as.double(ADM_DISC),
      RATE                 = as.double(RATE),
      INTERVAL_LOWER_LIMIT = as.double(INTERVAL_LOWER_LIMIT),
      INTERVAL_HIGHER_LIMIT= as.double(INTERVAL_HIGHER_LIMIT)) |> 
  sbt(!cheapr::is_na(RATE)) |> 
  roworder(HOSP_ID)
resp

cheapr::overview(resp)
```
